Tampering Detection In Low-Power \\ Smart Cameras

### 2015-06-22 paper.tex
final edits

We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras. The algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme: any frame yielding an outlier is detected as tampered. Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to improve the detection of camera displacements. Experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity. %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
\keywords{tampering detection, smart cameras, displacement detection, blurring detection}
\end{abstract}

@@

Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on simple indicators such as the average intensity (the \emph{luma}, denoted by $l$), the average frame difference (denoted by $d$) and the average norm of the gradient (denoted by $g$). The first two are meant to detect camera displacements, which would substantially change the image content, while the latter the blurring, which would cause a drop of the energy in the high-frequencies. As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$. The employed segmentation is detailed in Section~\ref{subsec:Segmentation}.

For each $z_t$, we compute the luma and frame difference (FD) separately on each of the $K$ regions,
\begin{eqnarray}
\vspace{-0.5cm}
l_k(t) &=& \frac{1}{\#R_k} \sum_{x \in R_k} z_t(x) \label{eq:lumaRegions}\\
d_k(t) &=&\frac{1}{\#R_k} \sum_{x \in R_k} \left(z_t(x) - z_{t-1}(x) \right)^2, \ \ k = 1, \dots, K\,,\label{eq:FDRegions}
\vspace{-0.5cm}
\end{eqnarray}
where $\#(\cdot)$ denotes the cardinality of a set. We can simultaneously monitor both luma and frame difference, even though computing $d_k$ requires to store the previous frame and this also depends on memory availability of the device. In what follows, including Algorithm~\ref{alg:DISPL}, we consider only the luma $l$, but the same procedures apply to the frame difference $d$ as well. The average gradient norm is instead computed over the whole image
\begin{equation}
\label{eq:normaGradiente}
g(t) = \sum_{x \in X} \left (\sqrt{\left(z_t \circledast f_h\right)^2(x) + \left(z_t \circledast f_v\right)^2(x)}\right),
\end{equation}
where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, and $\circledast$ denotes the 2d convolution.

\begin{figure}[tb]
\centering
\includegraphics[width=1\linewidth]{Immagini/indicatori}
\caption{Example of camera displacements. 
Left plots report the luma values, while right plots depict the detrended luma: 
we show both indicators computed over the full image (\textbf{a}) and on two reported regions (\textbf{b} and \textbf{c}). Red dots indicate a single displaced frame. Values in the highlighted areas refer to the frames depicted on top of the figure. 
Displacement yields a peak in $l$ and two outliers in the sequence of detrended indicators $\partial l$, which can be clearly detected.}
\label{fig:indicatori} 
\end{figure}
%Tampering changes the sequence of indicators:
Figure~\ref{fig:indicatori} shows how a camera displacement affects the luma. First of all, we observe that a displaced frame changes $l(t)$ (introducing a peak in Figure~\ref{fig:indicatori}(a)) and that this change is also visible in the indicators $l_k(t)$ for the regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)). Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here. In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables, while here indicators follow an unpredictable trend because of changes in the scene or in the illumination. Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative
\begin{equation}\label{eq:detrending}
 \partial l_k(t) = l_k(t)-l_k(t-1),  \ \ k = 1, \dots, K\,,
\end{equation}
and we similarly define $\partial d_k(t)$ and $\partial g(t)$.

@@

It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, a single tampered frame yields two outliers (as shown in Figure~\ref{fig:indicatori}) and only the first and the last of a sequence of consecutive tampered frames yield outliers in the detrended indicators. This is the reason why detrended indicators are monitored in a one-shot manner, targeting the detection of the first tampered-frame. On the one hand, this one-shot solution can provide prompt detections, which is particularly important at the low frame-rates we consider, on the other hand, the persistence of tampering is disregarded. To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, possibly combined with the proposed one-shot monitoring.
%
%
\subsection{Scene Segmentation}\label{subsec:Segmentation}
Scene has to be preliminarily segmented to define regions that Algorithm \ref{alg:DISPL} takes as input. To this purpose, we use part of the training frames before $T_0$ to compute the feature vector $\textbf{f}(x)\in \mathbb{R}^5$ for each pixel $x\in\mathcal{X}$ 
\begin{equation}
\label{eq:featureVector}
\textbf{f}(x)=\left[r(x);c(x);\bar{l}(x);\sigma_{l}(x);\overline{g}(x);\sigma_{g}(x)\right]\,, \ \forall x \in X\,.
\end{equation}
In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the mean and standard deviation of the intensity at $x$, computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ the mean of gradient norm and its standard deviation, computed over time. These feature vectors are meant to cluster pixels in regions having, over training frames, similar spatial appearance and temporal behavior.

As in superpixel methods \cite{Susstrunk2012}, segmentation is performed by k-means clustering. Feature vectors~\eqref{eq:featureVector} over the whole image are clustered by a weighted k-means~\cite{kottke1994motion} that scales each component of the Euclidean distance between a feature vector and a cluster centroid of a weight that is inversely proportional to the standard deviation over the cluster. This scaling compensates the fact that feature vector components might span very different ranges. The number of clusters is defined by testing several values and then choosing the best solution according to the Calinski-Harabasz criterion~\cite{calinski1974dendrite}.

@@

The most computationally demanding operations of Algorithm \ref{alg:DISPL} consist in computing the indicators, in particular $g$. Computing $g$ requires, when using the Sobel filters in~\eqref{eq:normaGradiente}, $34$ operations per pixels\footnote{Execution can be accelerated whether hardware implementations of the FFT transform are available}, while computing $l$ and $d$ requires 1 and 3 operations per pixel, respectively. Detecting outliers in the indicators requires two comparisons per frame: thus, monitoring regions instead of the whole image has a negligible impact on the overall computational complexity. Algorithm~\ref{alg:DISPL} has very low memory requirements: $l$ and $g$ indicators can be computed online, while $d$ requires to store a single frame in memory. Segmentation is executed only during the initial configuration and can be performed on an external device connected to the smart camera. As such, Algorithm~\ref{alg:DISPL} can be properly executed on low-power and ultra low-power cameras.
%\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
\vspace{-0.2cm}
\section{Experiments}\label{sec:experiments}
Experiments are meant to assess the advantages of separately monitoring regions in the considered tampering-detection framework. To this purpose, we show that Algorithm~\ref{alg:DISPL} detects camera displacements better than monitoring indicators on the whole image (\emph{Whole Image} in Figure~\ref{fig:ROC}). We also show that it is important to define regions considering the scene content, since operating on $K$ regions obtained by clustering $[r(x), c(x)]$ leads to performance loss (\emph{Voronoi} in Figure~\ref{fig:ROC}). For both Voronoi and regions defined as in Section~\ref{subsec:Segmentation}, we consider the two extreme configurations where alerts are raised at the first region firing an outlier or when $K-1$ regions simultaneously fire outliers.

We recorded $8$ sequences from webcams monitoring different urban areas, yielding overall $12200$ frames. From each frame we have cropped the central area, removing the 50 top-most and bottom-most rows, and the 50 left-most and right-most columns\footnote{Sequences can be provided upon request.}. We have synthetically introduced $10\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and coulumns (as in Figure \ref{fig:sequences}(b)). Blurring  (as in Figure \ref{fig:sequences}(a)) was simulated by convolution against a Gaussian kernel, with $1\leq\sigma\leq5$ randomly defined. 
%\gi{Adriano: image sizes? blur kernel std? quanti pixel il displacement?}
%The second dataset was recorded using a Raspberry Pi Model B+ \cite{raspberry} mounting a camera module. 
%Few sequences of the first dataset contain real tampering events  (Figure \ref{fig:indicatori}). 
%In the second dataset, tampering was introduced by moving the device or spraying water on the camera lens.

A typical figure of merit for assessing detection performance is the Receiver Operating Characteristic (ROC) curve, where each point corresponds to a pair (FPR$_\gamma$, TPR$_\gamma$) defined as
%
\[\text{FPR}_\gamma = \frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}, \ \text{ and } \ \text{TPR}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
%
Here, TP$_\gamma$ represents the number of true positives (tampered frames correctly detected), FP$_\gamma$ of false positives (tampering-free frames detected), TN$_\gamma$ of true negatives (tampering-free frames not detected) and FN$_\gamma$ of false negatives (missed tampered frames), for a given value of $\gamma$. ROC curves have been computed by varying $\gamma_l$, $\gamma_d$ and $\gamma_g$ in a range of values between $0.1$ and $50$. 
\vspace{-0.3cm}
\subsubsection{Discussion}
ROC curves in Figure~\ref{fig:ROC}(a)-(b) and the Area Under the Curve (AUC) values reported in the caption confirm that camera displacements can be better detected by monitoring FD than luma, and that it is convenient to define regions segmenting the scene content. As an example, if we set $\gamma_d = 15.6$, Algorithm~\ref{alg:DISPL} operates at $\text{FPR} = 1\%$ and detects nearly all tampering since $\text{TPR} = 99.92\%$. In contrast, to achieve $\text{FPR} = 1\%$ when monitoring the whole image, FPR drops to $91.67\%$ ($\gamma_d = 6.5$), and when considering Voronoi to $\text{TPR} = 89.05\%$ ($\gamma_d = 18.64$). When monitoring $l$, Algorithm~\ref{alg:DISPL} achieves TPR~=~$73.98\%$ at $\text{FPR}=1\%$  ($\gamma_d=5.8$), while monitoring the whole image is quite ineffective ($\text{TPR} = 17.85\%$, $\gamma_d=3.7$), and similarly perform Voronoi's regions ($\text{FPR} = 53.02\%$, $\gamma_d=6$). We remark that, in the considered low-power scenario, it is important to operate at low FPR, given the large number of frames that smart cameras have to process over time and the cost of data-transmission in terms of battery power. As far as the values of $\Gamma_d$ and $\Gamma_l$ are concerned, the most effective configuration consists in raising an alert as soon as a single region fires a very clear outlier, which implies setting quite wide intervals~\eqref{eq:confidenceRegions}. This option is also preferable because it enables the detection of partial occlusions of the scene.

%\gi{Aggiornare con i nuovi dati}
%\emph{Considering $d$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.4\%$ of false alarms, with $\gamma = 24.7$, which is less than one fifth of the whole image solution ($2.2\%$ with $\gamma = 4.6$) and the Voronoi solution ($2.3\%$ with $\gamma = 12.9$).
%On the other side, considering $l$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $13\%$ of false alarms, with $\gamma = 2.5$, which is about one sixth of the whole image solution ($75\%$ with $\gamma = 0.1$) and about three fifty than the Voronoi solution ($22\%$ with $\gamma = 1.7 $).
%Having few false alarms is important in a low-power context, considering the high number of frames that have to be processed. Most importantly, false alerts are halved basically at no computational overhead.} \gi{Aggiornare con i nuovi dati}
Figure~\ref{fig:ROC}(c) confirms that blurring is more effectively detected by monitoring the whole image at once. This is probably due to the fact that, typically, regions do not include the most prominent edge of the scene, which are indeed the most informative parts to detect blurring. This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.%\gi{aggiungere commento su K-1 vornoi che va meglio?}
x
\begin{figure}[t]
\centering
\includegraphics[width=1\linewidth]{Immagini/curve}
\caption{\textbf{(a)} ROC curves relative to displacement detection based on FD: the AUC obtained by 
``Algorithm 1 - One region'' is $99.89\%$, while the AUC obtained by on the whole 
image is $99.65\%$.
\textbf{(b)} ROC curves relative to displacement detection based on luma: the AUC obtained by 
``Algorithm 1 - One region'' is $98.44\%$, while the AUC obtained on the whole 
image is $84.07\%$.
\textbf{(c)} ROC curves relative to blurring detection based on gradient: as opposed to \textbf{(a)} and 
\textbf{(b)}, ``Algorithm 1 - One region'' (AUC $98.4\%$) is outperformed by monitoring the whole image, AUC = $99.13\%$.
To highlight differences, the ROC curves in \textbf{(a)} and \textbf{(c)} are drawn over a smaller FPR, TPR range.}
%\caption{\textbf{(a)} ROC curve for displacement detection considering Frame Difference (FD) as indicator.
%	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
%	The AUC corresponding to Algorithm~\ref{alg:DISPL} ($1$ region - Algorithm $1$) is $99.89\%$, while the AUC of considering the whole image (Total) is $99.65\%$. 
%	%the AUC of Luma-Voronoi is $99.85\%$, and the AUC of Frame-Difference is $99.44\%$ (axis have been zoomed for the visualization sake). Large AUC values indicate that these displacements can be successfully detected and that Algorithm~\ref{alg:DISPL} is the most effective one.
%	\textbf{(b)} ROC curve for displacement detection considering Luma as indicator.
%	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
%	The AUC corresponding to Algorithm~\ref{alg:DISPL} ($1$ region - Algorithm $1$) is $98.44\%$, while the AUC of considering the whole image (Total) is $84.07\%$. 
%	\textbf{(c)} ROC curve for blurring detection.
%	In this case, the Gradient-Total (AUC $99.13\%$) outperforms the Gradient-Regions (AUC $98.4\%$) because this latter do not include the most prominent edges in images, which are the most important information to detect blurring. %This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image. 
%	ROC curves \textbf{(a)} and \textbf{(c)}
%	%For the visualization sake, ROC curves \textbf{(a)} and \textbf{(c)} have been reported in a limited range of both axis, and in particular we focus on low values of 1-SPECIFICITY, thus of false alarms.
	%}
\label{fig:ROC}
\end{figure}
%\begin{figure}[htb]
%\centering
%\subfigure[]{\label{fig:buenosAiresDef}\includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus1}}
%\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus_fd}}
%\caption{ROC curves for defocus detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean gradient energy. \textbf{(b)} Analysis of the frame differencing.}
%\label{fig:ROCdisplacement}
%\label{fig:ROCdefocus}
%\end{figure}

@@

Ongoing work concerns approaching other types of tampering, such as degradations of the imaging sensor, and the integration of sequential monitoring schemes~\cite{alippi2010detecting} to detect subtle tampering that persists over time. Moreover, we will investigate the use of superpixels methods~\cite{Susstrunk2012} to perform the preliminary scene segmentation, by including suitable temporal information to the feature vectors.

>>>>

We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras. The algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme: any frame yielding an outlier is detected as tampered. Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to improve the detection of camera displacements. Experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity. %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
\keywords{tampering detection, smart cameras, displacement detection, blurring detection, low-power cameras, low-frame rate.}
\end{abstract}

@@

Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on simple indicators such as the average intensity (the \emph{luma}, denoted by $l$), the average frame difference (\emph{FD}, denoted by $d$) and the average norm of the gradient (denoted by $g$). The first two are meant to detect camera displacements, which would substantially change the image content, while the latter the blurring, which would attenuate the high frequencies of the image. As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$. The employed segmentation is detailed in Section~\ref{subsec:Segmentation}.

For each $z_t$, we compute the luma and frame difference separately on each of the $K$ regions,
\begin{eqnarray}
\vspace{-0.5cm}
l_k(t) &=& \frac{1}{\#R_k} \sum_{x \in R_k} z_t(x) \label{eq:lumaRegions}\\
d_k(t) &=&\frac{1}{\#R_k} \sum_{x \in R_k} \left(z_t(x) - z_{t-1}(x) \right)^2, \ \ k = 1, \dots, K\,,\label{eq:FDRegions}
\vspace{-0.5cm}
\end{eqnarray}
where $\#(\cdot)$ denotes the cardinality of a set. We can simultaneously monitor both luma and frame difference, even though computing $d_k$ requires to store the previous frame and this also depends on memory availability of the device. In what follows, including Algorithm~\ref{alg:DISPL}, we consider only the luma $l$, but the same procedures apply to the frame difference $d$ as well. The average gradient norm is instead computed over the whole image
\begin{equation}
\label{eq:normaGradiente}
g(t) = \sum_{x \in X} \left (\sqrt{\left(z_t \circledast f_h\right)^2(x) + \left(z_t \circledast f_v\right)^2(x)}\right),
\end{equation}
where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, and $\circledast$ denotes the 2d convolution.

\begin{figure}[tb]
\centering
\includegraphics[width=1\linewidth]{Immagini/indicatori}
\caption{Example of camera displacements. 
Left plots report the luma values, while right plots depict the detrended luma: 
we show both indicators computed over the full image (\textbf{a}) and on two reported regions (\textbf{b} and \textbf{c}). Red dots indicate a single displaced frame. Values in the highlighted areas refer to the frames depicted on top of the figure. 
The highlighted displacement yields a small peak in $l(t)$ and two outliers in the sequence of detrended indicators $\partial l(t)$, which can be clearly detected.}
\label{fig:indicatori} 
\end{figure}
%Tampering changes the sequence of indicators:
Figure~\ref{fig:indicatori} shows how a camera displacement affects the luma. First of all, we observe that a displaced frame changes $l(t)$ (introducing a peak in Figure~\ref{fig:indicatori}(a)) and that this change is also visible in the indicators $l_k(t)$ for the regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)). Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here. In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables, while here indicators follow an unpredictable trend because of changes in the scene or in the illumination. Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative
\begin{equation}\label{eq:detrending}
 \partial l_k(t) = l_k(t)-l_k(t-1),  \ \ k = 1, \dots, K\,,
\end{equation}
and we similarly define $\partial d_k(t)$ and $\partial g(t)$.

@@

It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, a single tampered frame yields two outliers (as shown in Figure~\ref{fig:indicatori}) and only the first and the last of a sequence of consecutive tampered frames yield outliers in the detrended indicators. This is the reason why detrended indicators are monitored in a \emph{one-shot} manner, targeting the detection of the first tampered-frame. On the one hand, this one-shot monitoring can provide prompt detections, which is particularly important at the low frame-rates we consider, on the other hand, the persistence of tampering is disregarded. To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, possibly combined with Algorithm~\ref{alg:DISPL}.
%
%
\subsection{Scene Segmentation}\label{subsec:Segmentation}
Scene has to be preliminarily segmented to define regions that Algorithm \ref{alg:DISPL} takes as input. To this purpose, we use part of the training frames before $T_0$ to compute the feature vector $\textbf{f}(x)\in \mathbb{R}^5$ for each pixel $x\in\mathcal{X}$ 
\begin{equation}
\label{eq:featureVector}
\textbf{f}(x)=\left[r(x);c(x);\bar{l}(x);\sigma_{l}(x);\overline{g}(x);\sigma_{g}(x)\right]\,, \ \forall x \in X\,.
\end{equation}
In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the mean and standard deviation of the intensity at $x$, computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ the mean of gradient norm and its standard deviation, computed over time. These feature vectors are meant to cluster pixels in regions having, over training frames, similar spatial appearance and temporal behavior.

As in superpixel methods \cite{Susstrunk2012}, segmentation is performed by k-means clustering. Feature vectors~\eqref{eq:featureVector} over the whole image are clustered by a weighted k-means~\cite{kottke1994motion} that scales each component of the Euclidean distance between a feature vector and a cluster centroid of a weight that is inversely proportional to the standard deviation over the cluster. This scaling compensates the fact that the components of the feature vector might span very different ranges. The number of clusters is defined by testing several values and then choosing the best solution according to the Calinski-Harabasz criterion~\cite{calinski1974dendrite}.

@@

The most computationally demanding operations of Algorithm \ref{alg:DISPL} consist in computing the indicators, in particular $g$. Computing $g$ requires, when using the Sobel filters in~\eqref{eq:normaGradiente}, $34$ operations per pixels\footnote{Execution can be accelerated whether hardware implementations of the FFT transform are available}, while computing $l$ and $d$ requires 1 and 3 operations per pixel, respectively. Detecting outliers in the indicators requires two comparisons per frame: thus, monitoring regions instead of the whole image has a negligible impact on the overall computational complexity. Algorithm~\ref{alg:DISPL} has very low memory requirements: $l$ and $g$ indicators can be computed online, while $d$ requires to store a single frame in memory. Segmentation is executed only during the initial configuration and can be performed on an external device connected to the smart camera. As such, Algorithm~\ref{alg:DISPL} can be properly executed on low-power and ultra low-power cameras.
%\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
\vspace{-0.2cm}
\section{Experiments}\label{sec:experiments}
Experiments are meant to assess the advantages of separately monitoring regions in the considered tampering-detection framework. To this purpose, we show that Algorithm~\ref{alg:DISPL} detects camera displacements better than monitoring indicators on the whole image (\emph{Whole Image} in Figure~\ref{fig:ROC}). We also show that it is important to adapt regions to the image content, since operating on $K$ regions obtained by clustering $[r(x), c(x)]$ leads to performance loss (\emph{Voronoi} in Figure~\ref{fig:ROC}). For both Voronoi and regions defined as in Section~\ref{subsec:Segmentation}, we consider the two extreme configurations where alerts are raised at the first region firing an outlier or when $K-1$ regions simultaneously fire outliers.

We recorded $8$ sequences from webcams monitoring different urban areas, yielding overall $12200$ frames. From each frame, we have cropped the central area, removing the 50 top-most and bottom-most rows, and the 50 left-most and right-most columns\footnote{Sequences can be provided upon request.}. We have synthetically introduced $10\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and columns (as in Figure \ref{fig:sequences}(b)), while blurring (as in Figure \ref{fig:sequences}(a)) was simulated by convolution against a Gaussian kernel, with standard deviation randomly defined in the range $[1,5]$. 
%\gi{Adriano: image sizes? blur kernel std? quanti pixel il displacement?}
%The second dataset was recorded using a Raspberry Pi Model B+ \cite{raspberry} mounting a camera module. 
%Few sequences of the first dataset contain real tampering events  (Figure \ref{fig:indicatori}). 
%In the second dataset, tampering was introduced by moving the device or spraying water on the camera lens.

A typical figure of merit for assessing detection performance is the Receiver Operating Characteristic (ROC) curve, where each point corresponds to a pair (FPR$_\gamma$, TPR$_\gamma$) defined as
%
\[\text{FPR}_\gamma = \frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}, \ \text{ and } \ \text{TPR}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
%
Here, TP$_\gamma$ represents the number of true positives (tampered frames correctly detected), FP$_\gamma$ of false positives (tampering-free frames detected), TN$_\gamma$ of true negatives (tampering-free frames not detected) and FN$_\gamma$ of false negatives (missed tampered frames), for a given value of $\gamma$. ROC curves have been computed by varying $\gamma_l$, $\gamma_d$ and $\gamma_g$ in a range of values between $0.1$ and $50$. 
\vspace{-0.3cm}
\subsubsection{Discussion}
ROC curves in Figure~\ref{fig:ROC}(a)-(b) and the Area Under the Curve (AUC) values reported in the caption confirm that camera displacements can be better detected by monitoring FD than luma, and that it is convenient to define regions by segmenting the scene. As an example, if we set $\gamma_d = 15.6$, Algorithm~\ref{alg:DISPL} operates at $\text{FPR} = 1\%$ and detects nearly all tampering since $\text{TPR} = 99.92\%$. In contrast, to operate at $\text{FPR} = 1\%$ when monitoring the whole image or Voronoi regions we obtain $\text{TPR} = 91.67\%$ ($\gamma_d = 6.5$) and $\text{TPR} = 89.05\%$ ($\gamma_d = 18.64$), respectively. When monitoring $l$, Algorithm~\ref{alg:DISPL} achieves TPR~=~$73.98\%$ at $\text{FPR}=1\%$  ($\gamma_d=5.8$), while monitoring the whole image is quite ineffective ($\text{TPR} = 17.85\%$, $\gamma_d=3.7$), and similarly perform Voronoi's regions ($\text{TPR} = 53.02\%$, $\gamma_d=6$). We remark that, in the considered low-power scenario, it is important to operate at low FPR to prevent useless data transmission (that would reduce the battery time-life) and the activation of eventual countermeasures. As far as the values of $\Gamma_d$ and $\Gamma_l$ are concerned, the most effective configuration consists in raising an alert as soon as a single region fires a very clear outlier, which implies setting quite wide intervals~\eqref{eq:confidenceRegions}. This option is also preferable because it enables the detection of partial occlusions of the scene. We observe that displacements typically change all the Voronoi regions, and this is why Voronoi outperforms Algorithm~\ref{alg:DISPL} in the configuration $K-1$ in both luma and FD.

Figure~\ref{fig:ROC}(c) confirms that blurring is more effectively detected by monitoring the whole image at once. This is probably due to the fact that, typically, regions do not include the most prominent edge of the scene, which are indeed the most informative parts to detect blurring. % This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.%\gi{aggiungere commento su K-1 vornoi che va meglio?}

\begin{figure}[t]
\centering
\includegraphics[width=1\linewidth]{Immagini/curve}
\caption{\textbf{(a)} ROC curves relative to displacement detection based on FD: the AUC obtained by 
``Algorithm 1 - One region'' is $99.89\%$, while the AUC obtained by on the whole 
image is $99.65\%$.
\textbf{(b)} ROC curves relative to displacement detection based on luma: the AUC obtained by 
``Algorithm 1 - One region'' is $98.44\%$, while the AUC obtained on the whole 
image is $84.07\%$.
\textbf{(c)} ROC curves relative to blurring detection based on gradient: as opposed to \textbf{(a)} and 
\textbf{(b)}, ``Algorithm 1 - One region'' (AUC $98.4\%$) is outperformed by monitoring the whole image, AUC = $99.13\%$.
To highlight differences, the ROC curves in \textbf{(a)} and \textbf{(c)} are drawn over a smaller FPR, TPR range.}
%\caption{\textbf{(a)} ROC curve for displacement detection considering Frame Difference (FD) as indicator.
%	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
%	The AUC corresponding to Algorithm~\ref{alg:DISPL} ($1$ region - Algorithm $1$) is $99.89\%$, while the AUC of considering the whole image (Total) is $99.65\%$. 
%	%the AUC of Luma-Voronoi is $99.85\%$, and the AUC of Frame-Difference is $99.44\%$ (axis have been zoomed for the visualization sake). Large AUC values indicate that these displacements can be successfully detected and that Algorithm~\ref{alg:DISPL} is the most effective one.
%	\textbf{(b)} ROC curve for displacement detection considering Luma as indicator.
%	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
%	The AUC corresponding to Algorithm~\ref{alg:DISPL} ($1$ region - Algorithm $1$) is $98.44\%$, while the AUC of considering the whole image (Total) is $84.07\%$. 
%	\textbf{(c)} ROC curve for blurring detection.
%	In this case, the Gradient-Total (AUC $99.13\%$) outperforms the Gradient-Regions (AUC $98.4\%$) because this latter do not include the most prominent edges in images, which are the most important information to detect blurring. %This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image. 
%	ROC curves \textbf{(a)} and \textbf{(c)}
%	%For the visualization sake, ROC curves \textbf{(a)} and \textbf{(c)} have been reported in a limited range of both axis, and in particular we focus on low values of 1-SPECIFICITY, thus of false alarms.
	%}
\label{fig:ROC}
\end{figure}
%\begin{figure}[htb]
%\centering
%\subfigure[]{\label{fig:buenosAiresDef}\includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus1}}
%\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus_fd}}
%\caption{ROC curves for defocus detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean gradient energy. \textbf{(b)} Analysis of the frame differencing.}
%\label{fig:ROCdisplacement}
%\label{fig:ROCdefocus}
%\end{figure}

@@

Ongoing work concerns approaching other types of tampering, such as degradations of the imaging sensor, and the integration of sequential monitoring schemes~\cite{alippi2010detecting} to detect subtle tampering that persists over time. Moreover, we will investigate the use of superpixels methods~\cite{Susstrunk2012} to segment the scene, by including suitable temporal information to the feature vectors.
